Learning From Massive Noisy Labeled Data for Visual Classification

### 2014-11-14 paper.tex
Nearly final

%%%%%%%%% TITLE
\title{Learning From Massive Weakly Labeled Data for Visual Classification}

@@

%%%%%%%%% ABSTRACT
\begin{abstract}
Large-scale strongly supervised datasets are crucial to train deep neural networks for various computer vision problems. Obtaining such well labeled data is usually expensive, but collecting web data with noisy labels is much more effortless. In this paper we study how to train a deep neural network to classify objects with a small set of clean annotations and massive noisy labeled data. A probabilistic model is proposed to describe the relationship between image data, noisy label and ground truth. We further integrate the model into a deep learning framework. Experiments are conducted on a collected large-scale web dataset. The results show that noisy labeled data benefit the training of deep neural networks, and exploiting our label noise model can further improve the performance.
\end{abstract}

@@

Deep learning with massive amount of supervised training data has recently shown very impressive improvement on multiple image recognition challenges including image classification~\cite{krizhevsky2012imagenet}, attribute learning~\cite{zhang2013panda}, scene classification~\cite{farabet2013learning}, etc. While state-of-the-art results have been continuously reported~\cite{zeiler2013visualizing,simonyan2014very,szegedy2014going}, all these methods requires reliable annotations from millions of images~\cite{deng2009imagenet} which are often expensive and time-consuming to obtain~\cite{deng2009imagenet}, preventing deep models from being quickly trained on new image recognition problems and making it necessary to develop new efficient labeling and training frameworks for deep learning.

One possible solution is to automatically collect large amount of annotations from the Internet web images (i.e. extracting tags from the surrounding texts or keywords from search engines) and directly use them as ground truth to train deep models. Unfortunately, these labels are extremely unreliable due to the various types of noise (i.e. labeling mistakes from annotators or computing errors from extraction algorithms). Many works have shown that these noisy labels could adversely impact the classification accuracy of the induced classifiers [3, 13, 15]. Various noise-robust algorithms are developed but experiments show that performances of classifiers inferred by robust algorithms are still affected by label noise. Other noise cleaning algorithms are developed [1, 3, ], but these approaches are difficult in distinguishing informative hard examples from harmful mislabeled ones.

Although annotating all the noisy data is costly, it is often easy to obtain a small amount of clean labels. Based on the observation of transferability of deep neural networks, people first initialize the parameters with a model pretrained on a larger yet related dataset [5], and then finetuning on the smaller dataset of specific task. Such methods may better avoid overfitting and utilize the relationships between two datasets. However, we find that training a deep model directly with massive noisy data could achieve better performance than finetuning from a pretrained ImageNet model over a small subset, especially when the relations between the two datasets are weak. Other approaches address the classification problem as semi-supervised learning where noisy labels are discarded [6]. These algorithms usually suffer from model complexity thus cannot be applied on large-scale datasets.

@@

Our goal here is to build an end-to-end deep learning system that is capable of training with both limited clean labels and massive amount of noisy labels more effectively. Figure~\ref{fig:overview} shows the general framework of our system. The major challenge of using noisy labeled data is to identify bad samples and keep them from drifting the model training. Intuitively, we have two types of evidence to help make decisions. One comes from the ground truth labels of other similar images, and the other comes from how confusing the input image itself. In this paper, we model these two types of information as latent variables and use them to bridge the semantic gap between our observed image and weak label. Additionally, we introduce two types of label noises explicitly, pure random noise and confusion noise to model the noise more directly.

\begin{itemize}
   \item \textbf{Pure random noise} occurs when a weak label has no semantic relationship with its image, for example, the samples in the upper red box of Figure~\ref{fig:overview}. This kind of noise is often caused by either the mismatch between web image and surrounding texts, or false conversion from text to label. The noisy label in this case is not acceptable.
   \item \textbf{Confusing noise} occurs when a weak label is a reasonable object class for its image other than the ground truth one. This kind of noise happens regularly in some sort of images. For example, the knitwear in the lower orange box of Figure~\ref{fig:overview} can be often misclassified as a sweater. However, in real applications, confusion among several related labels is generally acceptable to users.
\end{itemize}

We propose a novel probabilistic model to capture the relations among all these factors. We solve the problem by Expectation-Maximization (EM) algorithm and integrate it into a deep learning framework. To validate our approach, we collect a clothing classification dataset consisting of around $72,000$ clean labeled and one million noisy labeled data. Experiments on this dataset demonstrate that massive weakly labeled data can benefit the training of deep neural networks even further.

Our work is inspired by [7], which also modifies the deep neural networks to train effectively with noisy labeled data. However, their modification is simply implemented by adding a linear layer on top of the softmax layer, which builds on the assumption that label noise is independent of input images. We generalize [7] to a new deep learning framework by allowing the noise is related with input images. We find that by modeling the noise more appropriately, the deep models can benefit even more from noisy labels.

Our main contribution in this paper is a novel end-to-end deep learning framework that models the noisy web data generation on the Internet. The deep learning system is designed to learn to predict noise types, ground truth labels as well as noisy labels. The system is able to learn from a small set of clean annotations with a massive amount of noisy labels and find the best trade-off between using all the data.

@@

% Visual classification, which aims at categorizing objects into predefined semantic groups, is a traditional yet fundamental problem in computer vision. Starting from the pioneering work in~\cite{krizhevsky2012imagenet}, researchers have developed various deep learning approaches to solve object classification problem. While state-of-the-art results have been continuously reported~\cite{zeiler2013visualizing,simonyan2014very,szegedy2014going}, all the methods rely on large-scale strongly supervised training dataset~\cite{deng2009imagenet}, which consumes a lot of manpower to label millions of images. In many real-world applications, we can only get limited well labeled data. Training a deep neural network is thus difficult even if we finetune it from a pretrained ImageNet model. However, we can easily get substantial weakly supervised data by crawling images from the Internet and extracting keywords from their surrounding text. For example, product photos on \emph{Amazon} and \emph{Taobao} usually have relatively reasonable title and tags, which can be converted to weak object class and attributes. Thus it drives us to find some effective ways to do visual recognition by utilizing these massive information.

% Without losing generality, we focus on the task of clothes recognition, \ie, determining whether an image is about a T-shirt, a sweater, or a dress, \etc. This specific object classification problem draws much attention to many e-commerce companies, since it is a basic building block of products retrieval and recommendation systems.  The major challenge of using weakly labeled data in training classifier is to identify bad samples and keep them from drifting the model. Figure~\ref{fig:overview} shows some examples of label noise. Existing methods try to solve this problem from two aspects. One way is to model the relationship between true label and noisy label, but the model is too simple to fit in real-world applications. The other one exploits semi-supervised learning algorithms (such as label propagation), but their complexities are usually quadratic with the number of samples, thus cannot be applied on large-scale datasets.

% To conquer this problem, we first study why a weak label goes wrong and conclude two types of label noise:
% \begin{itemize}
%    \item \textbf{Pointless noise} occurs when a weak label has no semantic relationship with its image, for example, the samples in the upper red box of Figure~\ref{fig:overview}. This kind of noise is often caused by either the mismatch between web image and surrounding text, or false conversion from text to label. The noisy label in this case is not acceptable.
%    \item \textbf{Confusing noise} occurs when a weak label is a reasonable object class for its image other than the ground truth one. This kind of noise happens regularly in some sort of images. For example, the knitwear in the lower orange box of Figure~\ref{fig:overview} can be often misclassified as a sweater. However, in real applications, confusion among several related labels is generally acceptable to users.
% \end{itemize}

% Our goal is to identify the ground truth label given an image as well as its weak label, and then use the ground truth label as supervision instead of the weak one. Intuitively, we have two kinds of evidence to help us make decision. One comes from the ground truth labels of other similar images, and the other one comes from how confusing is the input image itself. In this paper, we model these two kinds of information as latent variables and use them to bridge the semantic gap between our observed image and weak label. A novel probabilistic model is proposed to build relations among all these factors. We solve the problem by Expectation-Maximization (EM) algorithm and integrate it into a deep learning framework. To validate our approach, we collected a clothes image dataset consisting of about $72,000$ strongly labeled and one million weakly labeled data. Experiments on this dataset show that massive weakly labeled data benefit the training of deep neural networks, and our approach can further improve the performance.

% Our contribution comes from two aspects. First, we investigate the cause of weak label in real-world data, and formulate the label noise in a probabilistic model. Second, we integrate the proposed model into a deep learning framework to help it resist bad supervision in training with weakly labeled data.

@@

\section{Related Work} % (fold)
\label{sec:related_work}

\noindent
\textbf{Semi-supervised learning:} Apart from modeling label noise explicitly, some semi-supervised learning algorithms are developed to utilize weakly labeled or even unlabeled data. Label Propagation method~\cite{zhu2002learning} explicitly uses ground truths of well labeled data to classify unlabeled samples. However, it suffers from computing pairwise distance, which has quadratic complexity with the number of data samples and cannot be applied on large-scale datasets. Weston \etal~\cite{weston2012deep} proposed to embed a pairwise loss in the middle layer of a deep neural network, which benefits the learning of discriminative features. But they need extra information about whether a pair of unlabeled images belong to same class, which cannot be obtained in our problem.

\noindent
\textbf{Transfer learning:} The success of Convolutional Neural Networks (CNNs) lies in their capability of learning rich and hierarchical image features. However, the model parameters cannot be properly learned when training data is not enough. Researchers proposed to conquer this problem by first initializing the CNN parameters with a model pretrained on a larger yet related dataset, and then finetuning the CNN on the smaller dataset of specific task~\cite{krizhevsky2012imagenet,oquab2013learning,azizpour2014generic}. Nevertheless, this kind of transfer learning scheme could be suboptimal when the two tasks are just loosely related. In our case of clothes recognition, we show that training a deep model directly with massive weakly label data achieves better performance compared with finetuning from a ImageNet pretrained model.

\noindent
\textbf{Noise Modeling with deep learning:} Various methods have been proposed to handle label noise in different problem settings, but there are very few works about deep learning from noisy labels~\cite{mnih2012learning,sukhbaatar2014learning,vahdat2013handling}. ~\cite{mnih2012learning} builds a simple noisy model for aerial images but only consider binary classification. [9] assumes label noise is independent from the true class label which is a simple and a specific case. [7] generalizes from [9 11] by considering multi-class classification and modeling class dependent noise, but they assume the noise is independent from source images, ignoring the hardness of labeling different images.

% Various methods have been proposed to handle label noise in different problem settings. Our work is inspired by~\cite{sukhbaatar2014learning}, where they linearly map the probability distribution of the true label to that of the noisy label. But their assumption --- noisy label and input image are conditionally independent given the true label --- is too strong. It cannot figure out whether an observed label is correct or not. Our proposed method overcomes this problem by first estimating how likely an image suffers from pointless or confusing noise. And then predicting the true label based on this information and our observed noisy label.

% Other methods model noisy labels in different context, \eg, visual attributes annotation~\cite{vahdat2013handling}, or roads recognition in aerial images~\cite{mnih2012learning}. It is non-trivial to generalize their approaches to solver our problem, \ie, multi-class object classification.

% Apart from modeling label noise explicitly, some semi-supervised learning algorithms are developed to utilize weakly labeled or even unlabeled data. Label Propagation method~\cite{zhu2002learning} explicitly uses ground truths of well-labeled data to classify unlabeled samples. However, it suffers from computing pairwise distance, which has quadratic complexity with the number of data samples and cannot be applied on large-scale datasets. Weston \etal~\cite{weston2012deep} proposed to embed a pairwise loss in the middle layer of a deep neural network, which benefits the learning of discriminative features. But they need extra information about whether a pair of unlabeled images belong to same class, which cannot be obtained in our problem.

% The success of Convolutional Neural Networks (CNNs) lies in their capability of learning rich and hierarchical image features. However, the model parameters cannot be properly learned when training data is not enough. Researchers proposed to conquer this problem by first initializing the CNN parameters with a model pretrained on a larger yet related dataset, and then finetuning the CNN on the smaller dataset of specific task~\cite{krizhevsky2012imagenet,oquab2013learning,azizpour2014generic}. Nevertheless, this kind of transfer learning scheme could be suboptimal when the two tasks are just loosely related. In our case of clothes recognition, we show that training a deep model directly with massive weakly label data achieves better performance compared with finetuning from a ImageNet pretrained model.

>>>>

%%%%%%%%% TITLE
\title{Learning From Massive Noisy Labeled Data for Visual Classification}

@@

%%%%%%%%% ABSTRACT
\begin{abstract}
Large-scale strongly supervised datasets are crucial to train deep neural networks for various computer vision problems. Obtaining such well labeled data is usually expensive, but collecting web data with noisy labels is much more effortless. In this paper we model label noise with a graphical model and further integrate it into a deep learning framework. Experiments are conducted on a collected large-scale web dataset. The results show that noisy labeled data benefit the training of deep neural networks, and exploiting our label noise model can further improve the performance.
\end{abstract}

@@

Deep learning with massive amount of supervised training data has recently shown very impressive improvement on multiple image recognition challenges including image classification~\cite{krizhevsky2012imagenet}, attribute learning~\cite{zhang2013panda}, scene classification~\cite{farabet2013learning}, etc. While state-of-the-art results have been continuously reported~\cite{zeiler2013visualizing,simonyan2014very,szegedy2014going}, all these methods require reliable annotations from millions of images~\cite{deng2009imagenet} which are often expensive and time-consuming to obtain~\cite{deng2009imagenet}, preventing deep models from being quickly trained on new image recognition problems. Thus it is necessary to develop new efficient labeling and training frameworks for deep learning.

One possible solution is to automatically collect large amount of annotations from the Internet web images~\cite{gong2014multiview} (i.e. extracting tags from the surrounding texts or keywords from search engines) and directly use them as ground truth to train deep models. Unfortunately, these labels are extremely unreliable due to the various types of noise (i.e. labeling mistakes from annotators or computing errors from extraction algorithms). Many works have shown that these noisy labels could adversely impact the classification accuracy of the induced classifiers~\cite{zhu2004class,nettleton2010study,pechenizkiy2006class}. Various noise-robust algorithms are developed but experiments show that performances of classifiers inferred by robust algorithms are still affected by label noise. Other noise cleaning algorithms are developed~\cite{barandela2000decontamination,brodley2011identifying}, but these approaches are difficult in distinguishing informative hard examples from harmful mislabeled ones.

Although annotating all the noisy data is costly, it is often easy to obtain a small amount of clean labels. Based on the observation of transferability of deep neural networks, people first initialize the parameters with a model pretrained on a larger yet related dataset~\cite{krizhevsky2012imagenet}, and then finetuning on the smaller dataset of specific task~\cite{oquab2013learning,azizpour2014generic}. Such methods may better avoid overfitting and utilize the relationships between the two datasets. However, we find that training a deep model directly with massive noisy data could achieve better performance than finetuning from a pretrained ImageNet model over a small subset, especially when the relations between the two datasets are weak. Other approaches address the classification problem as semi-supervised learning where noisy labels are discarded~\cite{zhu2002learning}. These algorithms usually suffer from model complexity thus cannot be applied on large-scale datasets.

@@

Our goal here is to build an end-to-end deep learning system that is capable of training with both limited clean labels and massive amount of noisy labels more effectively. Figure~\ref{fig:overview} shows the general framework of our system. The major challenge of using noisy labeled data is to identify bad samples and keep them from drifting the model training. Intuitively, we have two types of evidence to help make decisions. One comes from the ground truth labels of other similar images, and the other comes from how confusing the input image itself. In this paper, we model these two types of information as latent variables and use them to bridge the semantic gap between our observed image and weak label. Additionally, we introduce two types of label noise:
\begin{itemize}
   \item \textbf{Pure random noise} occurs when a noisy label has no semantic relationship with its image, for example, the samples with ``$\times$'' in Figure~\ref{fig:overview}. This kind of noise is often caused by either the mismatch between web image and surrounding texts, or false conversion from text to label. The noisy label in this case is not acceptable.
   \item \textbf{Confusing noise} occurs when a noisy label is a reasonable object class for its image other than the ground truth one. This kind of noise happens regularly in some sort of images. For example, the samples with ``?'' in Figure~\ref{fig:overview}. However, in real applications, confusion among several related labels is generally acceptable to users.
\end{itemize}

We propose a novel probabilistic model to capture the relations among all these factors. We solve the problem by Expectation-Maximization (EM) algorithm and integrate it into a deep learning framework. To validate our approach, we collect a clothing classification dataset consisting of around $72,000$ clean labeled and one million noisy labeled data. Experiments on this dataset demonstrate that massive noisy labeled data could benefit the training of deep neural networks.

Our work is inspired by~\cite{sukhbaatar2014learning}, which also modifies the deep neural networks to train effectively with noisy labeled data. However, their modification is simply implemented by adding a linear layer on top of the softmax layer, which builds on the assumption that label noise is independent with input images. We generalize~\cite{sukhbaatar2014learning} to a new deep learning framework by relating the noise to the input image. We find that by modeling the noise more appropriately, the deep models can benefit even more from noisy labels.

@@

@@

\section{Related Work} % (fold)
\label{sec:related_work}
For most of the related works including effect of label noises, taxonomy of label noises, robust algorithms and noise cleaning algorithms for learning with noisy data, we refer to \cite{frenay2013classification} for a comprehensive review.

{\bf Direct learning with noisy labels:} Among other consequences, many works have shown that noise can adversely impact the classification performances of induced classifiers \cite{zhu2004class}. 
Experiments in the literature show that the performances of classifiers inferred by label noise-robust algo- rithms are still affected by label noise. Label noise-robust methods seem to be adequate only for simple cases of label noise that can be safely managed by overfitting avoidance \cite{frenay2013classification}.

{\bf Semi-supervised learning:} Apart from modeling label noise explicitly, some semi-supervised learning algorithms are developed to utilize weakly labeled or even unlabeled data. Label Propagation method \cite{zhu2002learning} explicitly uses ground truths of well-labeled data to classify unlabeled samples. However, it suffers from computing pairwise distance, which has quadratic complexity with the number of data samples and cannot be applied on largescale datasets. Weston et al. \cite{weston2012deep} proposed to embed a pairwise loss in the middle layer of a deep neural net- work, which benefits the learning of discriminative features. But they need extra information about whether a pair of unlabeled images belong to same class, which cannot be obtained in our problem.

{\bf Transfer learning:} The success of Convolutional Neural Networks (CNNs) lies in their capability of learning rich and hierarchical image features. However, the model parameters cannot be properly learned when training data is not enough. Researchers proposed to conquer this problem by first initializing the CNN parameters with a model pretrained on a larger yet related dataset, and then finetuning the CNN on the smaller dataset of specific task \cite{krizhevsky2012imagenet, oquab2013learning, azizpour2014generic}. Nevertheless, this kind of transfer learning scheme could be suboptimal when the two tasks are just loosely related. In our case of clothes recognition, we show that training a deep model directly with massive weakly label data achieves bette performance compared with finetuning from a ImageNet pretrained model.

{\bf Noise modeling with deep learning:} Various methods have been proposed to handle label noise in different problem settings, but there are very few works about deep learning from noisy labels \cite{mnih2012learning, larsen1998design, sukhbaatar2014learning}. \cite{mnih2012learning} builds a simple noisy model for aerial images but only consider binary classification. \cite{larsen1998design} assumes label noise is independent from the true class label which is a simple and a specific case. \cite{sukhbaatar2014learning} generalizes from them by considering multi-class classification and modeling class dependent noise, but they assume the noise is independent from source images, ignoring the hardness of labeling different images.
% section related_work (end)
